# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r}
data <- read.csv("activity.csv")

head(data)
nrow(data)
nrow(subset(data, is.na(steps)))
```

## What is mean total number of steps taken per day?

```{r}
steps_per_day <- aggregate(steps ~ date, data, sum)
hist(steps_per_day$steps,
     main = "Histogram of the total number of steps taken each day",
     xlab = "Steps per Day",
     ylab = "Frequency")
```

```{r}
mean1 <- mean(steps_per_day$steps)
```

The mean is ```r as.character(mean1)``` steps per day.

```{r}
median1 <- median(steps_per_day$steps)
```

The median is ```r as.character(median1)``` steps per day.

## What is the average daily activity pattern?

```{r}
steps_per_interval <- aggregate(steps ~ interval, data, mean)
summary(steps_per_interval)
plot(steps_per_interval$steps ~ steps_per_interval$interval,
     type = "l",
     main = "Average daily activity pattern",
     xlab = "Interval",
     ylab = "Average Steps")
```

```{r}
max_steps <- max(steps_per_interval$steps)
tmp <- subset(steps_per_interval, steps == max_steps)
tmp$interval
```

5-minute interval which contains the maximum number of steps is ```r tmp$interval```.


## Imputing missing values

```{r}
tmp <- nrow(subset(data, is.na(steps)))
```

The total number of missing values in the dataset is ```r tmp```.

```{r}
data2 <- data
steps_per_interval2 <- aggregate(steps ~ interval, data2, mean)

for (i in 1:nrow(data2))
{
    # If the steps column is NA,
    # fill it with average value.
    if (is.na(data2[i,c("steps")]))
    {
        # interval
        j = data2[i,3]

        # fetch a record from the average data set,
        # and fill an empty "steps" with it.
        avg_row <- subset(steps_per_interval2, interval == j)
        avg_steps <- avg_row[1,2]
        data2[i,1] <- avg_steps
    }
}

nrow(data2)

tmp <- nrow(subset(data2, is.na(steps)))
```

Now, the total number of missing values in the dataset is ```r tmp```.

```{r}
steps_per_day2 <- aggregate(steps ~ date, data2, sum)
hist(steps_per_day2$steps,
     main = "Histogram of the total number of steps taken each day",
     xlab = "Steps per Day",
     ylab = "Frequency")
```

```{r}
mean2 <- mean(steps_per_day2$steps)
```

The mean is ```r as.character(mean2)``` steps per day.

```{r}
median2 <- median(steps_per_day2$steps)
```

The median is ```r as.character(median2)``` steps per day.

So, here is the differences.

According to the histgrams, the frequency of the 10000-15000 class is increased.

And

```{r}
mean_diff <- mean2 - mean1
```

the mean is differed ```r mean_diff``` steps from the first one.

```{r}
median_diff <- median2 - median1
```

Also the median is differed ```r median_diff``` steps from the first one.


## Are there differences in activity patterns between weekdays and weekends?

```{r}
# changing locale stuff because running Rstudio in Japanese environment
Sys.setlocale(category = "LC_TIME", locale = "C")
Sys.setlocale(category = "LC_ALL", locale = "C")

for (i in 1:nrow(data2))
{
    if (weekdays(as.Date(data2[i,c("date")])) %in% c("Saturday", "Sunday"))
    {
        data2[i,c("weekday")] <- "weekend"
    }
    else
    {
        data2[i,c("weekday")] <- "weekday"
    }
}

data_weekday <- subset(data2, weekday == "weekday")
data_weekend <- subset(data2, weekday == "weekend")

par(mfrow = c(2,1), ps = 11, mar = c(4,4,2,2))

steps_per_interval_weekday <- aggregate(steps ~ interval, data_weekday, mean)
plot(steps_per_interval_weekday$steps ~ steps_per_interval_weekday$interval,
     type = "l",
     main = "weekday",
     xlab = "Interval",
     ylab = "Average Steps")

steps_per_interval_weekend <- aggregate(steps ~ interval, data_weekend, mean)
plot(steps_per_interval_weekend$steps ~ steps_per_interval_weekend$interval,
     type = "l",
     main = "weekend",
     xlab = "Interval",
     ylab = "Average Steps")
```
